Bayesian spectral density estimation using P-splines with quantile-based knot placement

### 2020-07-01 pSplines.tex
some edits

\title{Bayesian spectral density estimation using P-splines}
\author{Patricio Maturana-Russel \and Renate Meyer}

@@

This article  proposes a Bayesian approach to estimating the spectral density of a stationary time series using a prior based on a mixture of P-spline distributions.  Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2019} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations. The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots. Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients. We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves similar ability to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.

@@

Bayesian nonparametric prior specifications for the psd include \cite{Cart:1997}
who used a smoothing prior on the log psd.  \cite{Rosen:2012} decompose the log psd into a linear and nonlinear component and  put a Gaussian process prior on the nonlinear part, \cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
%
\cite{Gangopadhyay:1999} modelled the log-spectral density function by  piecewise polynomials of low order between strategically placed knots on the support of the psd,  allowing the data to select the number and location of the knots.  A reversible jump Markov chain Monte Carlo \citep[RJMCMC;][]{Green:1995} algorithm was implemented to sample from the posterior distribution.  Similarly, frequentist approaches such as \cite{Rodriguez} and \cite{Wood2017} model the log spectral density by a linear combination of splines with variable knots and combine these with adaptive smoothing.

Penalized splines have been used for a time-domain analysis of time series data e.g.\ to model the mean as a smooth function of time by \cite{Krivobokova}  or for fitting and  forecasting univariate  nonlinear time series by \cite{Wegener}, but not in the frequency domain for estimating the spectral density.
The P-spline prior on the spectral density function that we are going to propose is related to the 
Bernstein-Dirichlet process (BDP) prior  introduced by \cite{Choudhuri:2004} and combined with the Whittle likelihood.  The BDP prior  is a mixture of beta densities with weights induced by a Dirichlet process.  The number of components, which is a smoothing parameter, is given a discrete prior.  \cite{Edwards2019} extended the method by replacing the beta densities with B-spline densities. Similar to 
\cite{Gangopadhyay:1999}, the specification of the number and location of the B-spline densities is data-driven, however it avoids a RJMCMC algorithm by putting a second Dirichlet process prior  on the cdf that induces the inter-knot distances. 
Unlike the beta densities  which have global support on the interval [0,1], B-spline densities have local support, thereby increasing their flexibility and allowing better modelling of spectral densities with sharp peaks and abrupt changes.  As demonstrated in \cite{Edwards2019}, the B-spline--Dirichlet process (BspDP) prior outperforms the BDP prior in estimating complex psd's.  However, this flexibility comes at a high computational price.

@@

$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``Adaptive knots".
%Therefore, the P-spline prior on the spectral density function is given by
%\begin{align*}
%f(\pi \omega) = \tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1],
%\end{align*}
%where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant.

@@

%\subsection*{Knot allocation strategy}
\subsection*{Adaptive knots}

The knots are often equally spaced in P-spline methods.  This scheme works quite well  for smooth psds without any abrupt changes.  However, for spectral densities with spikes, the algorithm will require a large number of knots to adequately estimate the peaks.  The B-spline prior method is able to handle abrupt changes because the knot location is variable and  driven by the data.  We propose an adaptive method that allocates the knots according to the periodogram but once allocated, the knots stay fixed.  The idea is to concentrate more effort in those regions that have potential peaks as detected by the periodogram.

Our adaptive knot location proposal works as follow.  We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.  Second, we standardize these values and apply the absolute value function.  We treat this transformed periodogram as a density function $\widetilde{f}$ and calculate its distribution function $F$.  Finally, the knots are allocated according to the quantiles of the inverse distribution function of $\widetilde{f}$.  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes. Hereafter, we will refer to this scheme as {\em adaptive knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
%REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?

The procedure can be summarized as follows:
\begin{enumerate}
	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ equally spaced values in $[0,1]$.
	\item The adaptive knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.

@@

The adaptive knot allocation can be regarded as an empirical Bayes prior since the data are used to specify the prior distribution.  However, only the locations of the B-splines are determined by the data, not the parameters of the prior distribution which in this case are the weights $w_k$ associated with each B-spline $b_{k,r}$.  The main criticism of an empirical Bayes approach is that the data are being used twice, once to specify the prior and then again to update the prior to the posterior. Here, the data are used only once to specify the knot locations but not again to update these through the likelihood.

@@

The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd} \citep{psplinepackage}.  A total of 100,000 MCMC samples were generated
using  a pilot run of 20,000 iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.  The final sample consisted of 80,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,500 samples used for posterior inferences.  
%Therefore, the number of iterations equate the B-spline analysis.  
This procedure was replicated for both equidistant and adaptive knots considering first and second order penalties.

@@

For the AR(1) time series, the P-spline prior with equidistant knots and penalty order $d=1$ yields better psd estimates than the BspDP prior.
A larger penalty order does not necessarily result in more accurate psd estimates. In this case of a relatively smooth spectral density, the adaptive knots do not yield an improvement over the equidistant knots and the BspDP prior.

@@

For the AR(4) time series, the fit based on the P-spline prior with adaptive knots is better than the fit based on the B-spline Dirichlet process prior and the one based on the P-spline prior with equidistant knots.
% categorically outperforms the BspDP algorithm.   It also outperforms the equidistant knot scheme. 
 In general, the BspDP algorithm achieves higher accuracy than the P-spline method with equidistant knots.  As shown by \cite{Edwards2019}, the posterior distribution based on the BspDP prior yields better results than posterior based on the BDP in terms of median IAE in this example.  Consequently, our P-spline algorithm with adaptive knots outperforms the BDP method as well.

@@

%REMARK: PLEASE INSERT GRAPH WITH AR(1) AND AR(4) PSD ESTIMATES
\def\arraystretch{1.1}
% IAE
\begin{table}
	\centering
%	\begin{ruledtabular}
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &  0.825 & 0.769 & 0.609  \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  0.772 & 0.644 & 0.527 \\
			\hspace{0.5em}P-spline $d=2$&  0.708 & 0.604 & 0.629 \\
			\textit{Adaptive knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 0.874 & 0.783 & 0.610 \\
			\hspace{0.5em}P-spline $d=2$ & 0.874 & 0.815 & 0.640 \\ \hline
			AR(4)     &  &  &  \\ 
			B-spline & 3.019 & 2.250 & 1.832   \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  2.914 & 2.389 & 2.307\\
			\hspace{0.5em}P-spline $d=2$&  3.150 & 2.569 & 2.388\\
			\textit{Adaptive knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 2.495 & 1.830 & 1.489\\
			\hspace{0.5em}P-spline $d=2$ &  2.696 & 2.003 & 1.677\\
			\bottomrule
		\end{tabular}
%	\end{ruledtabular}
	\caption{Median IAE for the psd estimates using the BspDP and the P-spline priors for simulated AR(1) and AR(4) time series with  different penalty orders ($d$) and knot location schemes for the P-splines.}
	\label{table:sim_IAE}
\end{table}

@@

For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in 
almost all the 300 independent analyses.  In general, the latter method shows diminishing  performance as the penalty degree increases.  For the AR(4) model, the P-spline method with equidistant knots performs poorly.  On the other hand, the adaptive knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.  %For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.  
For the P-spline priors, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.  As shown by \cite{Edwards2019}, the BspDP algorithm produces better results than the BDP prior in the AR(4) case, which coverage probability is zero across $n=\{128,256,512\}$. Consequently, our P-spline algorithm with adaptive knots outperforms the BDP method in this case.

Even though the P-spline method with equidistant knots for the AR(4) case performs poorly in terms of uniform coverage, i.e. the proportion of times that the credible band covers  the true psd {\em completely} in the 300 runs, it is important to note that this is only because the uniform band does not cover
the psd for  just very few frequencies. As is evident from 
Table~\ref{table:sim_covPbb_median},  only for a small fraction of the frequencies the psd is not covered by the uniform credible band.  In this context, BspDP and adaptive knot P-spline algorithms perform equally well.

@@

% Coverage probability Table
\begin{table}
	\centering
%	\begin{ruledtabular}
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &  1.000 & 1.000 & 1.000 \\
			\textit{Equidistant}& & &   \\
			\hspace{0.5em}P-spline $d=1$&  1.000 & 1.000 & 0.997\\
			\hspace{0.5em}P-spline $d=2$& 0.993 & 1.000 & 0.993  \\
			\textit{Adaptive knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 0.997 & 1.000 & 0.997 \\
			\hspace{0.5em}P-spline $d=2$ & 0.973 & 0.740 & 0.770 \\ \hline		       
			AR(4)     &  &  &  \\ 
			B-spline &  0.933 & 0.967 & 0.897  \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  0.527 & 0.353 & 0.007 \\
			\hspace{0.5em}P-spline $d=2$ & 0.323 & 0.167 & 0.007\\
			\textit{Adaptive knots}& & &   \\	
			\hspace{0.5em}P-spline $d=1$ & 0.790 & 0.847 & 0.813 \\
			\hspace{0.5em}P-spline $d=2$ & 0.337 & 0.387 & 0.270 \\
			\bottomrule
		\end{tabular}
%	\end{ruledtabular}
	\caption{Proportion of times that the true psd is entirely encapsulated within the 90\% uniform credible band of the 300 estimates.}
	\label{table:sim_covPbb}
\end{table}

\begin{table}
	\centering
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &  1.000 & 1.000 & 1.000 \\
			\textit{Equidistant \&}& & &   \\
			\textit{Adaptive knots}& & &   \\
			\hspace{0.5em}P-spline $d=1,2$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
			AR(4)     &  &  &  \\ 
			B-spline &  1.000 & 1.000 & 1.000 \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  1.000 & 0.992 & 0.984 \\
			\hspace{0.5em}P-spline $d=2$&  0.984 & 0.984 & 0.980 \\
			\textit{Adaptive knots}& & &   \\	
			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
			\hspace{0.5em}P-spline $d=2$ & 0.984 & 0.992 & 0.992\\	
			\bottomrule
		\end{tabular}
	\caption{Median of the pointwise  proportions that the true psd at each Fourier frequency is contained within the $90\%$ uniform credible bands.}
	\label{table:sim_covPbb_median}
\end{table}

@@

As \cite{Edwards2019} noted, one of the drawbacks of the BspDP prior  is its computational complexity relative to the BDP.  This is directly reflected in the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2019}).  The median run-time in our analyses for the BspDP prior and the P-spline with adaptive knots are displayed in Table~\ref{table:sim_time}.  Note that the P-spline method based on equidistant knots has similar run-times to the adaptive knots.  The P-spline method is approximately 4-7 times faster than BspDP approach in these examples.  This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.

% TIME Table
\begin{table}
	\centering
%	\begin{ruledtabular}
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &   29.497 & 32.423 & 37.896 \\
			\textit{Adaptive knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$&  4.859 & 7.450 & 9.606\\
			\hspace{0.5em}P-spline $d=2$&  4.538 & 6.863 & 8.783\\
			B-spline/P-spline &  6.278 & 4.531 & 4.122 \\ \hline
			AR(4)     &  &  & \\ 
			B-spline & 30.111 & 34.126 & 40.875  \\
			\textit{Adaptive knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 4.883 & 7.453 & 9.665 \\
			\hspace{0.5em}P-spline $d=2$& 4.543 & 6.887 & 8.827 \\
			B-spline/P-spline & 6.389 & 4.760 & 4.421 \\	
			\bottomrule
		\end{tabular}
%	\end{ruledtabular}
	\caption{Median run-times in minutes and relative run-times with respect to an average of the P-spline run-times.}
	\label{table:sim_time}
\end{table}

In general terms, the P-spline algorithm with our novel knot allocation scheme, the adaptive knots, offers the best trade-off between accuracy and computation cost.  The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.

@@

The MCMC algorithm based on the P-spline prior is run in two stages.  In  the first stage, the proposal  is calibrated for the final run.  It consists of 25,000 iterations with a burn-in period of 10,000 and thinning factor of 10, resulting in 1,500 samples.  Then the algorithm is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples.  The analysis is performed using the first order penalty for equidistant and adaptive knots.   All other specifications are the same as those used in the simulation study.  A single analysis takes about 4.5 minutes.

\begin{figure}[]
	\centering
	\includegraphics[scale=0.40,clip=true,angle=0]{sunspot.pdf}
	\caption{Log-spectral density estimate for the sunspot data. The continuous grey line represents the periodogram.  The continuous and dashed lines stand for the posterior median log psd obtained using the B-spline and P-spline priors, respectively.  The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and adaptive} knots.
	\label{fig:sunspot}
\end{figure}

>>>>

\title{Bayesian spectral density estimation using P-splines with quantile-based knot placement}
\author{Patricio Maturana-Russel \and Renate Meyer}

@@

This article  proposes a Bayesian approach to estimating the spectral density of a stationary time series using a prior based on a mixture of P-spline distributions.  Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2019} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations. The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.  \textcolor{red}{In addition to equidistant knots, a novel strategy for a more expedient placement of  knots is proposed that makes use of the information provided by the periodogram about the steepness of the spectral power distribution.}
We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves similar ability to accurately estimate peaks due to the new data-driven knot allocation scheme but significantly reduces the computational costs.

@@

Bayesian nonparametric prior specifications for the psd include \cite{Cart:1997}
who used a smoothing prior on the log psd.  \cite{Rosen:2012} segmented a nonstationary time series into stationary segments. \textcolor{red}{For each stationary segment, they decomposed the log psd into a linear and nonlinear component and  put a linear smoothig spline prior on the nonlinear part. }\cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
%
\cite{Gangopadhyay:1999} modelled the log-spectral density function by  piecewise polynomials of low order between strategically placed knots on the support of the psd,  allowing the data to select the number and location of the knots.  A reversible jump Markov chain Monte Carlo \citep[RJMCMC;][]{Green:1995} algorithm was implemented to sample from the posterior distribution.  Similarly, frequentist approaches such as \cite{Rodriguez} and \cite{Wood2017} model the log spectral density by a linear combination of splines with variable knots and combine these with adaptive smoothing.

Penalized splines have been used for a time-domain analysis of time series data e.g.\ to model the mean as a smooth function of time by \cite{Krivobokova}  or for fitting and  forecasting univariate  nonlinear time series by \cite{Wegener}, but not in the frequency domain for estimating the spectral density.
The P-spline prior on the spectral density function that we are going to propose is related to the 
Bernstein-Dirichlet process (BDP) prior  introduced by \cite{Choudhuri:2004} and combined with the Whittle likelihood.  The BDP prior  is a mixture of beta densities with weights induced by a Dirichlet process.  The number of components, which is a smoothing parameter, is given a discrete prior.  \cite{Edwards2019} extended the method by replacing the beta densities with B-spline densities. Similar to 
\cite{Gangopadhyay:1999}, the specification of the number and location of the B-spline densities is data-driven, however it avoids a RJMCMC algorithm by putting a second Dirichlet process prior  on the cdf that induces the inter-knot distances. \textcolor{red}{ \cite{Choudhuri:2004} already demonstrated in a comprehensive simulation study  that the Bayesian psd estimates based on the BDP prior outperforms in terms of the $L_1$-error the smoothed periodogram estimates based on Bartlett-Priestley quadratic kernel, the penalized MLE of \cite{Pawitan} and autoregressive spectral estimates in most scenarios but noted that the BDP-based estimates detect peaks correctly but often underestimate the magnitude of sharp peaks.}
Unlike the beta densities  which have global support on the interval [0,1], B-spline densities have local support, thereby increasing their flexibility and allowing better modelling of spectral densities with sharp peaks and abrupt changes.  As demonstrated in \cite{Edwards2019}, the B-spline--Dirichlet process (BspDP) prior outperforms the BDP prior in estimating complex psd's.  However, this flexibility comes at a high computational price.

@@

$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. \textcolor{red}{A more judicious scheme for the placement  of knots that not only speeds up posterior computations but improves the fit for peaked spectral densities is described in  section ``Quantile-based knot placement".}
%Therefore, the P-spline prior on the spectral density function is given by
%\begin{align*}
%f(\pi \omega) = \tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1],
%\end{align*}
%where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant.

@@

%\subsection*{Knot allocation strategy}
\subsection*{Quantile-based knot placement}

The knots are often equally spaced in P-spline methods.  This scheme works quite well  for smooth psds without any abrupt changes.  However, for spectral densities with spikes, the algorithm will require a large number of knots to adequately estimate the peaks.  The B-spline prior method is able to handle abrupt changes because the knot location is variable and  driven by the data.  We propose a new selective method that allocates the knots according to quantiles of a periodogram-related distribution function but once allocated, the knots stay fixed.  The idea is to concentrate more effort in those regions that have potential peaks as detected by the periodogram.  \textcolor{red}{It is similar in spirit to the knot placement of O-splines \cite{WandOrmerod} in the regression context that make use of the empirical quantiles of the covariates. }

Our knot location proposal works as follow.  We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.  Second, we standardize these values and apply the absolute value function.  We treat this transformed periodogram as a density function $\widetilde{f}$ and calculate its distribution function $F$.  Finally, the knots are allocated according to the quantiles of the inverse distribution function of $\widetilde{f}$.  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes. Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
%REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?

The procedure can be summarized as follows:
\begin{enumerate}
	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.

@@

The quantile based knot allocation can be regarded as an empirical Bayes prior since the data are used to specify the prior distribution.  However, only the locations of the B-splines are determined by the data, not the parameters of the prior distribution which in this case are the weights $w_k$ associated with each B-spline $b_{k,r}$.  The main criticism of an empirical Bayes approach is that the data are being used twice, once to specify the prior and then again to update the prior to the posterior. Here, the data are used only once to specify the knot locations but not again to update these through the likelihood.

@@

The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd} \citep{psplinepackage}.  A total of 100,000 MCMC samples were generated
using  a pilot run of 20,000 iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.  The final sample consisted of 80,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,500 samples used for posterior inferences.  
%Therefore, the number of iterations equate the B-spline analysis.  
This procedure was replicated for both equidistant and Q-spaced knots considering first and second order penalties.

@@

For the AR(1) time series, the P-spline prior with equidistant knots and penalty order $d=1$ yields better psd estimates than the BspDP prior.
A larger penalty order does not necessarily result in more accurate psd estimates. In this case of a relatively smooth spectral density, the Q-spaced knots do not yield an improvement over the equidistant knots and the BspDP prior.

@@

For the AR(4) time series, the fit based on the P-spline prior with Q-spaced knots is better than the fit based on the B-spline Dirichlet process prior and the one based on the P-spline prior with equidistant knots.
% categorically outperforms the BspDP algorithm.   It also outperforms the equidistant knot scheme. 
 In general, the BspDP algorithm achieves higher accuracy than the P-spline method with equidistant knots.  As shown by \cite{Edwards2019}, the posterior distribution based on the BspDP prior yields better results than posterior based on the BDP in terms of median IAE in this example.  Consequently, our P-spline algorithm with Q-spaced knots outperforms the BDP method as well.

@@

%REMARK: PLEASE INSERT GRAPH WITH AR(1) AND AR(4) PSD ESTIMATES
\def\arraystretch{1.1}
% IAE
\begin{table}
	\centering
%	\begin{ruledtabular}
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &  0.825 & 0.769 & 0.609  \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  0.772 & 0.644 & 0.527 \\
			\hspace{0.5em}P-spline $d=2$&  0.708 & 0.604 & 0.629 \\
			\textit{Q-spaced knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 0.874 & 0.783 & 0.610 \\
			\hspace{0.5em}P-spline $d=2$ & 0.874 & 0.815 & 0.640 \\ \hline
			AR(4)     &  &  &  \\ 
			B-spline & 3.019 & 2.250 & 1.832   \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  2.914 & 2.389 & 2.307\\
			\hspace{0.5em}P-spline $d=2$&  3.150 & 2.569 & 2.388\\
			\textit{Q-spaced knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 2.495 & 1.830 & 1.489\\
			\hspace{0.5em}P-spline $d=2$ &  2.696 & 2.003 & 1.677\\
			\bottomrule
		\end{tabular}
%	\end{ruledtabular}
	\caption{Median IAE for the psd estimates using the BspDP and the P-spline priors for simulated AR(1) and AR(4) time series with  different penalty orders ($d$) and knot location schemes for the P-splines.}
	\label{table:sim_IAE}
\end{table}

@@

For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in 
almost all the 300 independent analyses.  In general, the latter method shows diminishing  performance as the penalty degree increases.  For the AR(4) model, the P-spline method with equidistant knots performs poorly.  On the other hand, the quantile-based knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.  %For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.  
For the P-spline priors, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.  As shown by \cite{Edwards2019}, the BspDP algorithm produces better results than the BDP prior in the AR(4) case, which coverage probability is zero across $n=\{128,256,512\}$. Consequently, our P-spline algorithm with Q-spaced knots outperforms the BDP method in this case.

Even though the P-spline method with equidistant knots for the AR(4) case performs poorly in terms of uniform coverage, i.e. the proportion of times that the credible band covers  the true psd {\em completely} in the 300 runs, it is important to note that this is only because the uniform band does not cover
the psd for  just very few frequencies. As is evident from 
Table~\ref{table:sim_covPbb_median},  only for a small fraction of the frequencies the psd is not covered by the uniform credible band.  In this context, BspDP and Q-spaced  knot P-spline algorithms perform equally well.

@@

% Coverage probability Table
\begin{table}
	\centering
%	\begin{ruledtabular}
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &  1.000 & 1.000 & 1.000 \\
			\textit{Equidistant}& & &   \\
			\hspace{0.5em}P-spline $d=1$&  1.000 & 1.000 & 0.997\\
			\hspace{0.5em}P-spline $d=2$& 0.993 & 1.000 & 0.993  \\
			\textit{Q-spaced knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 0.997 & 1.000 & 0.997 \\
			\hspace{0.5em}P-spline $d=2$ & 0.973 & 0.740 & 0.770 \\ \hline		       
			AR(4)     &  &  &  \\ 
			B-spline &  0.933 & 0.967 & 0.897  \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  0.527 & 0.353 & 0.007 \\
			\hspace{0.5em}P-spline $d=2$ & 0.323 & 0.167 & 0.007\\
			\textit{Q-spaced knots}& & &   \\	
			\hspace{0.5em}P-spline $d=1$ & 0.790 & 0.847 & 0.813 \\
			\hspace{0.5em}P-spline $d=2$ & 0.337 & 0.387 & 0.270 \\
			\bottomrule
		\end{tabular}
%	\end{ruledtabular}
	\caption{Proportion of times that the true psd is entirely encapsulated within the 90\% uniform credible band of the 300 estimates.}
	\label{table:sim_covPbb}
\end{table}

\begin{table}
	\centering
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &  1.000 & 1.000 & 1.000 \\
			\textit{Equidistant \&}& & &   \\
			\textit{Q-spaced knots}& & &   \\
			\hspace{0.5em}P-spline $d=1,2$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
			AR(4)     &  &  &  \\ 
			B-spline &  1.000 & 1.000 & 1.000 \\
			\textit{Equidistant knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ &  1.000 & 0.992 & 0.984 \\
			\hspace{0.5em}P-spline $d=2$&  0.984 & 0.984 & 0.980 \\
			\textit{Q-spaced knots}& & &   \\	
			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
			\hspace{0.5em}P-spline $d=2$ & 0.984 & 0.992 & 0.992\\	
			\bottomrule
		\end{tabular}
	\caption{Median of the pointwise  proportions that the true psd at each Fourier frequency is contained within the $90\%$ uniform credible bands.}
	\label{table:sim_covPbb_median}
\end{table}

@@

As \cite{Edwards2019} noted, one of the drawbacks of the BspDP prior  is its computational complexity relative to the BDP.  This is directly reflected in the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2019}).  The median run-time in our analyses for the BspDP prior and the P-spline with Q-spaced knots are displayed in Table~\ref{table:sim_time}.  Note that the P-spline method based on equidistant knots has similar run-times to the Q-spaced knots.  The P-spline method is approximately 4-7 times faster than BspDP approach in these examples.  This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.

% TIME Table
\begin{table}
	\centering
%	\begin{ruledtabular}
		\begin{tabular}{lccc}
			\toprule
			& $n=128$ & $n=256$ & $n=512$ \\ \hline
			AR(1)     &  &  &  \\ 
			B-spline &   29.497 & 32.423 & 37.896 \\
			\textit{Q-spaced knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$&  4.859 & 7.450 & 9.606\\
			\hspace{0.5em}P-spline $d=2$&  4.538 & 6.863 & 8.783\\
			B-spline/P-spline &  6.278 & 4.531 & 4.122 \\ \hline
			AR(4)     &  &  & \\ 
			B-spline & 30.111 & 34.126 & 40.875  \\
			\textit{Q-spaced knots}& & &   \\
			\hspace{0.5em}P-spline $d=1$ & 4.883 & 7.453 & 9.665 \\
			\hspace{0.5em}P-spline $d=2$& 4.543 & 6.887 & 8.827 \\
			B-spline/P-spline & 6.389 & 4.760 & 4.421 \\	
			\bottomrule
		\end{tabular}
%	\end{ruledtabular}
	\caption{Median run-times in minutes and relative run-times with respect to an average of the P-spline run-times.}
	\label{table:sim_time}
\end{table}

In general terms, the P-spline algorithm with our novel quantile-based  knot allocation scheme offers the best trade-off between accuracy and computation cost.  The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.

@@

The MCMC algorithm based on the P-spline prior is run in two stages.  In  the first stage, the proposal  is calibrated for the final run.  It consists of 25,000 iterations with a burn-in period of 10,000 and thinning factor of 10, resulting in 1,500 samples.  Then the algorithm is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples.  The analysis is performed using the first order penalty for equidistant and Q-spaced knots.   All other specifications are the same as those used in the simulation study.  A single analysis takes about 4.5 minutes.

\begin{figure}[]
	\centering
	\includegraphics[scale=0.40,clip=true,angle=0]{sunspot.pdf}
	\caption{Log-spectral density estimate for the sunspot data. The continuous grey line represents the periodogram.  The continuous and dashed lines stand for the posterior median log psd obtained using the B-spline and P-spline priors, respectively.  The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and Q-spaced knots.}
	\label{fig:sunspot}
\end{figure}
